%% Chapter 1: System Overview & Architecture
\chapter{System Overview \& Architecture}
\label{ch:overview}

\section{Introduction}

The \textbf{Aurxon Lab Access Management System (ALAMS)} is a production-grade, security-focused access control platform designed for educational computer laboratories. ALAMS replaces anonymous Windows desktop access with a verified, two-factor authentication system that proves \emph{who} a student is and \emph{which physical machine} they are sitting at.

\begin{tcolorbox}[alamsbox, title={\faInfoCircle\quad System Purpose}]
Every student who uses a lab computer is identified, authenticated, and tracked from the moment they sit down until they log out. No student may access the Windows desktop without completing the ALAMS authentication flow.
\end{tcolorbox}

\section{Core Objectives}

\begin{itemize}[leftmargin=*, label={\textcolor{aurxonblue}{\faArrowRight}}]
  \item \textbf{Identity Verification} — Know exactly who is using each machine at all times.
  \item \textbf{Physical Binding} — Prove a student is sitting at the specific computer they are unlocking.
  \item \textbf{Session Logging} — Record login time, logout time, and duration for every session.
  \item \textbf{Attendance Automation} — Generate classroom attendance records automatically.
  \item \textbf{Anonymous Access Prevention} — No student can access the desktop without authentication.
  \item \textbf{Security Enforcement} — Detect and respond to bypass attempts within 2 seconds.
  \item \textbf{Centralized Control} — Manage all lab computers from one dashboard.
\end{itemize}

\section{System Architecture}

ALAMS is composed of four main components that work together:

\begin{figure}[h!]
\centering
\begin{tikzpicture}[
  node distance=1.8cm and 2.5cm,
  box/.style={rectangle, rounded corners=8pt, minimum width=3cm, minimum height=1.2cm, text centered, font=\small\bfseries, draw, line width=1.2pt},
  arrow/.style={-{Stealth[length=6pt]}, line width=1.2pt}
]

%% Nodes
\node[box, fill=aurxondark!90, text=white, draw=aurxonblue] (server) at (0,0) {\faServer\\[2pt] Express Server\\{\tiny Node.js + WebSocket}};

\node[box, fill=aurxondark!90, text=white, draw=aurxongreen, below=of server] (db) {\faDatabase\\[2pt] Neon PostgreSQL\\{\tiny Cloud Database}};

\node[box, fill=aurxondark!90, text=white, draw=aurxonamber, right=3cm of server] (dashboard) {\faDesktop\\[2pt] Admin Dashboard\\{\tiny Next.js Web App}};

\node[box, fill=aurxondark!90, text=white, draw=aurxonred, left=3cm of server] (client) {\faLock\\[2pt] WPF Lock Client\\{\tiny C\# .NET 8}};

\node[box, fill=aurxondark!90, text=white, draw=aurxonblue!60, below=1.4cm of client] (watchdog) {\faShield*\\[2pt] Watchdog Service\\{\tiny NT SYSTEM}};

\node[box, fill=aurxondark!90, text=white, draw=aurxongreen!60, right=3cm of db] (mobile) {\faMobileAlt\\[2pt] Mobile Gateway\\{\tiny Student Browser}};

%% Arrows
\draw[arrow, aurxonblue] (client) -- node[above, font=\tiny, text=aurxonblue]{WebSocket} (server);
\draw[arrow, aurxonamber] (dashboard) -- node[above, font=\tiny, text=aurxonamber]{REST + WS} (server);
\draw[arrow, aurxongreen] (server) -- node[right, font=\tiny, text=aurxongreen]{Prisma ORM} (db);
\draw[arrow, aurxonred] (watchdog) -- node[right, font=\tiny, text=aurxonred]{HTTP POST} (client);
\draw[arrow, aurxonblue!60] (watchdog.east) to[bend right=30] node[below, font=\tiny, text=aurxonblue!60]{Heartbeat} (server.south);
\draw[arrow, aurxongreen!60] (mobile) -- node[above, font=\tiny, text=aurxongreen!60]{REST API} (server);

\end{tikzpicture}
\caption{ALAMS System Architecture — Component Interaction Diagram}
\label{fig:architecture}
\end{figure}

\subsection{Component 1: Express API Server}

The central backbone of ALAMS. Built with \textbf{Node.js}, \textbf{Express}, and \textbf{WebSocket (ws)}. Handles:
\begin{itemize}
  \item REST API endpoints for authentication, session management, reporting
  \item Real-time WebSocket connections from WPF client workstations
  \item Dynamic QR token generation and verification
  \item Admin control commands (remote lock/unlock)
  \item Watchdog heartbeat collection
\end{itemize}

\subsection{Component 2: Neon PostgreSQL Database}

Cloud-hosted relational database powered by \textbf{Neon}. Stores:
\begin{itemize}
  \item User accounts (students, admins, supervisors)
  \item Lab and computer registry
  \item Session and attendance records
  \item Security alerts and audit log
\end{itemize}

\subsection{Component 3: Admin Dashboard (Next.js)}

Web-based control center accessible from any browser. Features:
\begin{itemize}
  \item Live Monitor Grid with real-time workstation status
  \item Pending Computer Enrollment approval workflow
  \item Student directory management
  \item Faculty Attendance roster
  \item Pilot Analytics with live charts
  \item Security Audit log
\end{itemize}

\subsection{Component 4: WPF Lock Client + Watchdog Service}

Installed on every lab workstation:
\begin{itemize}
  \item \textbf{AlamsClient.exe} — Fullscreen WPF lock screen replacing the Windows shell (explorer.exe). Displays dynamic QR codes, accepts enrollment input, handles offline PIN login.
  \item \textbf{AlamsWatchdog.exe} — Windows service running as \texttt{NT AUTHORITY\textbackslash SYSTEM}. Monitors the lock client every 2 seconds. If the desktop shell runs without the ALAMS client, it immediately forces user logoff and logs a critical security alert.
\end{itemize}

\section{Authentication Flows}

ALAMS supports three authentication methods, each providing progressively lower security assurance:

\begin{table}[h!]
\centering
\caption{Authentication Method Comparison}
\label{tab:auth-methods}
\begin{tabularx}{\textwidth}{lXXl}
\toprule
\textbf{Method} & \textbf{How It Works} & \textbf{Use Case} & \textbf{Security Level}\\
\midrule
\rowcolor{aurxonlightgray}
QR Scan & Student scans dynamic QR on mobile, confirms on phone & Primary method & \textcolor{aurxongreen}{\textbf{HIGH}}\\
PIN Fallback & Student enters enrollment + 6-digit PIN locally & No smartphone available & \textcolor{aurxonamber}{\textbf{MEDIUM}}\\
\rowcolor{aurxonlightgray}
Admin Override & Administrator unlocks from dashboard & Emergency/demo & \textcolor{aurxonred}{\textbf{ADMIN ONLY}}\\
\bottomrule
\end{tabularx}
\end{table}

\section{Data Model Overview}

\begin{figure}[h!]
\centering
\begin{tikzpicture}[
  entity/.style={rectangle, rounded corners=4pt, draw=aurxonblue, fill=aurxonlightgray, minimum width=2.8cm, minimum height=0.8cm, font=\small\bfseries, text centered},
  rel/.style={-{Stealth[length=5pt]}, aurxongray, line width=0.8pt}
]
\node[entity] (user) at (0,0) {users};
\node[entity] (session) at (4,0) {sessions};
\node[entity] (attendance) at (8,0) {attendance};
\node[entity] (computer) at (4,-2.5) {computers};
\node[entity] (lab) at (0,-2.5) {labs};
\node[entity] (alert) at (8,-2.5) {security\_alerts};

\draw[rel] (user) -- node[above,font=\tiny]{1:N} (session);
\draw[rel] (session) -- node[above,font=\tiny]{1:N} (attendance);
\draw[rel] (computer) -- node[right,font=\tiny]{1:N} (session);
\draw[rel] (lab) -- node[above,font=\tiny]{1:N} (computer);
\draw[rel] (session) -- node[right,font=\tiny]{1:N} (alert);
\draw[rel] (computer) -- node[below,font=\tiny]{1:N} (alert);
\draw[rel] (user) -- node[right,font=\tiny]{1:N} (attendance);
\end{tikzpicture}
\caption{ALAMS Entity Relationship Diagram}
\label{fig:erd}
\end{figure}
