%% Chapter 5: Admin Dashboard & Mobile Gateway
\chapter{Web Dashboard \& Mobile Gateway}
\label{ch:dashboard}

The ALAMS web application serves two distinct interfaces from the same Next.js deployment: the \textbf{Admin Control Deck} for administrators/faculty, and the \textbf{Student Mobile Gateway} for QR-based authentication.

\section{Installing the Web Application}

\begin{tcolorbox}[codebox, title={Install Web App Dependencies}]
\begin{lstlisting}[language=bash]
cd "d:\Project Data Aurxon\ALAMS\web"
npm install
\end{lstlisting}
\end{tcolorbox}

\section{Configuring the Web Application}

Edit \texttt{web/.env.local} to point to your server:

\begin{tcolorbox}[codebox, title={web/.env.local}]
\begin{lstlisting}[language=bash]
# Replace SERVER_LAN_IP with the actual IP of the server machine
# Example: NEXT_PUBLIC_API_URL="http://192.168.1.100:5000"
NEXT_PUBLIC_API_URL="http://SERVER_LAN_IP:5000"
\end{lstlisting}
\end{tcolorbox}

\begin{tcolorbox}[warningbox, title={\faExclamationTriangle\quad Critical}]
The \envvar{NEXT\_PUBLIC\_API\_URL} must be the LAN IP of the server, not \texttt{localhost}. Student mobiles accessing the QR verification page cannot reach \texttt{localhost}; they need the network IP.
\end{tcolorbox}

\section{Starting the Dashboard}

\begin{tcolorbox}[codebox, title={Start the Next.js Web Application}]
\begin{lstlisting}[language=bash]
cd "d:\Project Data Aurxon\ALAMS\web"
npm run dev

# Accessible at:
# Admin Dashboard : http://SERVER_IP:3000/admin/login
# Student Mobile  : http://SERVER_IP:3000/login
# QR Unlock Page  : http://SERVER_IP:3000/unlock
\end{lstlisting}
\end{tcolorbox}

\section{Admin Login}

\begin{enumerate}[leftmargin=*, label=\textbf{\arabic*.}]
  \item Open a browser on the admin/faculty machine.
  \item Navigate to: \texttt{http://SERVER\_IP:3000/admin/login}
  \item Enter credentials:
    \begin{itemize}
      \item \textbf{Enrollment Number:} \texttt{ADMIN01}
      \item \textbf{Password:} \texttt{Admin@ALAMS2026!}
    \end{itemize}
  \item Click \textbf{Authenticate Access}.
  \item You are redirected to the Admin Control Deck.
\end{enumerate}

\section{Admin Dashboard Tabs}

\subsection{Live Monitor}
Real-time grid of all registered workstations. Each card shows:
\begin{itemize}[label={\textcolor{aurxonblue}{\faCircle[regular]}}]
  \item PC Number and Device Name
  \item Current Status (LOCKED / IN\_USE / PENDING / OFFLINE)
  \item Client Agent heartbeat status (ONLINE/OFFLINE — 15-second window)
  \item Watchdog Service status (ACTIVE/MISSING — 20-second window)
  \item Last contact timestamp
  \item Remote Lock / Remote Unlock buttons
\end{itemize}

\subsection{Pilot Analytics}
Computed KPIs from live server data:
\begin{itemize}[label={\textcolor{aurxongreen}{\faChartBar}}]
  \item \textbf{Login Success Rate} — Successful sessions vs. failed attempts (\%)
  \item \textbf{Average QR Unlock Time} — Computed from recorded \texttt{unlockLatencyMs}
  \item \textbf{Bypasses Blocked} — Watchdog enforcement events
  \item \textbf{Hourly Session Distribution} — Bar chart, last 24 hours (7:00–20:00)
  \item \textbf{Verification Method Breakdown} — QR / PIN / Admin ratios
\end{itemize}

\subsection{Faculty Attendance}
Live classroom roster for faculty:
\begin{itemize}[label={\textcolor{aurxonblue}{\faUserCheck}}]
  \item Search by student name or enrollment number
  \item Filter by lab zone
  \item Status: PRESENT (active) or DISCONNECTED (logged out)
\end{itemize}

\subsection{Session Audit}
Chronological audit trail of all sessions:
\begin{itemize}[label={\textcolor{aurxongray}{\faHistory}}]
  \item Session UUID, student identity, PC assignment
  \item Verification method badge: \textcolor{aurxonblue}{DYNAMIC QR} / \textcolor{aurxonamber}{PIN FALLBACK} / \textcolor{aurxonred}{ADMIN OVERRIDE}
  \item Login time, logout time, session status
\end{itemize}

\subsection{Labs \& Hardware}
\begin{itemize}[label={\textcolor{aurxonamber}{\faCog}}]
  \item Create new lab zones (name + location)
  \item Register workstations manually (PC number, device name, IP, MAC)
  \item Toggle PIN fallback per workstation
  \item Approve pending workstations (self-registered via WPF client)
\end{itemize}

\section{Student Mobile Gateway}

When a student scans the QR code on the lock screen, their mobile browser opens the ALAMS student gateway.

\subsection{Student Login Flow}

\begin{enumerate}[leftmargin=*, label=\textbf{Step \arabic*.}]
  \item Student mobile browser opens: \texttt{http://SERVER\_IP:3000/login}
  \item Student enters:
    \begin{itemize}
      \item \textbf{Enrollment Number:} e.g. \texttt{ENR2026001}
      \item \textbf{Password:} \texttt{Student@2026!}
    \end{itemize}
  \item After login, student is redirected to the QR verification page.
  \item The QR token from the lock screen URL is passed automatically.
  \item Student taps \textbf{Authorize Access}.
  \item Server verifies student identity + QR token + computer registration.
  \item Unlock signal is pushed via WebSocket to the workstation.
  \item Desktop unlocks within seconds.
\end{enumerate}

\section{Building for Production}

\begin{tcolorbox}[codebox, title={Production Build}]
\begin{lstlisting}[language=bash]
cd "d:\Project Data Aurxon\ALAMS\web"

# Build optimized production bundle
npm run build

# Start production server
npm start
# Serves at port 3000 (or set PORT env variable)
\end{lstlisting}
\end{tcolorbox}
