%% Chapter 10: Admin Daily Operations
\chapter{Admin Daily Operations}
\label{ch:operations}

\section{Starting the ALAMS System (Daily Startup)}

Follow this sequence each morning before the lab session begins:

\begin{enumerate}[leftmargin=*, label=\textbf{\arabic*.}, itemsep=6pt]
  \item \textbf{Start the API Server}
  \begin{tcolorbox}[codebox, title={Terminal 1 — API Server}]
  \begin{lstlisting}[language=bash]
  cd "d:\Project Data Aurxon\ALAMS\server"
  npm run dev
  # Or for production: pm2 start alams-server
  \end{lstlisting}
  \end{tcolorbox}

  \item \textbf{Start the Web Dashboard}
  \begin{tcolorbox}[codebox, title={Terminal 2 — Web Dashboard}]
  \begin{lstlisting}[language=bash]
  cd "d:\Project Data Aurxon\ALAMS\web"
  npm run dev
  # Or for production: npm start
  \end{lstlisting}
  \end{tcolorbox}

  \item \textbf{Log in to Admin Dashboard}\\
  Navigate to \texttt{http://SERVER\_IP:3000/admin/login}\\
  Enter: \credential{ADMIN01}{Admin@ALAMS2026!}

  \item \textbf{Verify all workstations are ONLINE}\\
  Check the Live Monitor tab — all registered PCs should show green ONLINE status within 30 seconds of the server starting.

  \item \textbf{Check for pending registrations}\\
  If any new workstations show as PENDING, approve them before the session starts.
\end{enumerate}

\section{Monitoring Live Sessions}

The \textbf{Live Monitor} tab auto-refreshes every 8 seconds. During a class session:

\begin{itemize}[label={\textcolor{aurxonblue}{\faEye}}]
  \item PCs in use show the \textcolor{aurxongreen}{\textbf{ACTIVE}} badge
  \item Locked but online PCs show \textcolor{aurxonamber}{\textbf{LOCKED}}
  \item Click \textbf{Refresh Deck} for an immediate status update
\end{itemize}

\subsection{Remote Lock (Kicking a Session)}

To forcibly terminate a student session from the dashboard:

\begin{enumerate}[leftmargin=*, label=\textbf{\arabic*.}]
  \item Find the workstation card in Live Monitor.
  \item Click the red \textbf{Force Lock} button.
  \item The student's desktop closes immediately; the lock screen returns.
  \item The session is recorded as TERMINATED in the audit log.
\end{enumerate}

\subsection{Admin Override (Emergency Unlock)}

To unlock a workstation without student authentication:

\begin{enumerate}[leftmargin=*, label=\textbf{\arabic*.}]
  \item Find the target workstation card.
  \item Click the green \textbf{Admin Override} button.
  \item The desktop unlocks under an ADMIN\_OVERRIDE session.
  \item This event is logged with the admin's identity and timestamp.
\end{enumerate}

\section{Faculty Attendance View}

Faculty can view real-time class attendance without admin credentials (using the Supervisor account):

\begin{enumerate}[leftmargin=*, label=\textbf{\arabic*.}]
  \item Log in with: \credential{FAC001}{Faculty@2026!}
  \item Click \textbf{Faculty Attendance} in the sidebar.
  \item See the live roster of all authenticated students.
  \item Use the search box to find specific students.
  \item Filter by lab zone if managing multiple rooms.
\end{enumerate}

\section{Generating Attendance Reports}

\subsection{Attendance Ledger Tab}

\begin{enumerate}[leftmargin=*, label=\textbf{\arabic*.}]
  \item Click \textbf{Attendance Ledger} in the sidebar.
  \item View all check-in/check-out records for today.
  \item Each row shows: Student Name, Enrollment Number, Lab, PC, Check-in time, Check-out time, Status.
\end{enumerate}

\subsection{Session Audit Tab}

Full audit trail including verification methods:
\begin{enumerate}[leftmargin=*, label=\textbf{\arabic*.}]
  \item Click \textbf{Session Audits} in the sidebar.
  \item View all sessions in reverse chronological order.
  \item Color-coded badges indicate authentication method.
  \item Terminated sessions are flagged for review.
\end{enumerate}

\section{Pilot Analytics Review}

After a lab session, review performance metrics:

\begin{enumerate}[leftmargin=*, label=\textbf{\arabic*.}]
  \item Click \textbf{Pilot Analytics} in the sidebar.
  \item Review:
    \begin{itemize}[label={--}]
      \item \textbf{Login Success Rate} — Should be $\geq 95\%$
      \item \textbf{Avg QR Unlock Time} — Target $\leq 10$ seconds
      \item \textbf{Bypasses Blocked} — Should be 0 in normal operation
    \end{itemize}
  \item The hourly chart shows peak usage periods.
\end{enumerate}

\section{End of Session Checklist}

At the end of each lab session:

\begin{itemize}[label={\textcolor{aurxongreen}{\faCheckSquare[regular]}}]
  \checkitem Verify all sessions have ended (no active students in Live Monitor)
  \checkitem Check Security Audits for any unresolved alerts
  \checkitem Export or screenshot the Attendance Ledger for records
  \checkitem Confirm all PCs have returned to LOCKED status
  \warnitem If a PC shows IN\_USE with no student present, use Remote Lock
\end{itemize}

\section{Resolving Security Alerts}

\begin{enumerate}[leftmargin=*, label=\textbf{\arabic*.}]
  \item Click \textbf{Security Audits} in the sidebar.
  \item Unresolved alerts show a yellow \textbf{Resolve} button.
  \item Investigate the alert (check the PC, talk to the student).
  \item Click \textbf{Resolve} to mark it as handled.
  \item All resolutions are logged with the admin ID and timestamp.
\end{enumerate}

\begin{table}[h!]
\centering
\caption{Security Alert Types}
\label{tab:alert-types}
\begin{tabularx}{\textwidth}{llXl}
\toprule
\textbf{Alert Type} & \textbf{Severity} & \textbf{Meaning} & \textbf{Action}\\
\midrule
\rowcolor{aurxonlightgray}
\texttt{failed\_login} & WARNING & Wrong PIN/credentials attempted & Check for unauthorized access\\
\texttt{watchdog\_kill} & CRITICAL & Student tried to bypass lock screen & Interview student; escalate\\
\rowcolor{aurxonlightgray}
\texttt{watchdog\_offline} & WARNING & Watchdog not reporting & Check service on that PC\\
\bottomrule
\end{tabularx}
\end{table}
