%% Chapter 8: Computer Registration & Enrollment
\chapter{Computer Registration \& Enrollment}
\label{ch:registration}

Every lab workstation must go through a one-time registration process before it can be used by students. This chapter describes the full registration workflow.

\section{Registration Workflow Overview}

\begin{figure}[h!]
\centering
\begin{tikzpicture}[
  step/.style={rectangle, rounded corners=6pt, draw=aurxonblue, fill=aurxonlightgray, minimum width=4cm, minimum height=0.9cm, text centered, font=\small\bfseries},
  admin/.style={rectangle, rounded corners=6pt, draw=aurxongreen, fill=green!10, minimum width=4cm, minimum height=0.9cm, text centered, font=\small\bfseries},
  arrow/.style={-{Stealth[length=6pt]}, line width=1pt, aurxongray}
]
\node[step] (s1) at (0,0) {1. MSI Installer Runs};
\node[step] (s2) at (0,-1.6) {2. First Boot — Auto-Register};
\node[step] (s3) at (0,-3.2) {3. PENDING State Displayed};
\node[admin] (s4) at (0,-4.8) {4. Admin Approves on Dashboard};
\node[step] (s5) at (0,-6.4) {5. QR Seed Provisioned via WebSocket};
\node[step] (s6) at (0,-8.0) {6. Client Enters QR LOCKED State};

\draw[arrow] (s1) -- (s2);
\draw[arrow] (s2) -- (s3);
\draw[arrow] (s3) -- (s4);
\draw[arrow] (s4) -- (s5);
\draw[arrow] (s5) -- (s6);

% Labels
\node[right=0.3cm of s2, font=\tiny, text=aurxongray] {SHA-256 fingerprint sent to server};
\node[right=0.3cm of s4, font=\tiny, text=aurxongreen] {Admin enters PC\# and Lab zone};
\node[right=0.3cm of s5, font=\tiny, text=aurxongray] {config.json updated on workstation};
\end{tikzpicture}
\caption{Workstation Registration Sequence}
\label{fig:registration-flow}
\end{figure}

\section{Step-by-Step Registration Process}

\subsection{Step 1: Run the MSI Installer}

\begin{tcolorbox}[codebox, title={Install ALAMS on Each Workstation}]
\begin{lstlisting}[language=PowerShell]
# Run as Administrator on each lab workstation
msiexec /i AlamsInstaller.msi /quiet /norestart

# The installer:
# - Copies AlamsClient.exe to C:\Program Files\ALAMS\
# - Copies AlamsWatchdog.exe to C:\Program Files\ALAMS\
# - Creates C:\ProgramData\ALAMS\ directory
# - Registers and starts the AlamsWatchdog service
\end{lstlisting}
\end{tcolorbox}

\subsection{Step 2: Log In as Student User}

Log out of the administrator account and log in as the dedicated student user. The ALAMS lock screen appears instead of the Windows desktop.

\subsection{Step 3: Automatic Hardware Fingerprint Registration}

When the client first runs:
\begin{enumerate}[leftmargin=*, label={\textcolor{aurxonblue}{\faCircle[regular]}}]
  \item Reads \texttt{MachineGuid} from \texttt{HKLM\textbackslash SOFTWARE\textbackslash Microsoft\textbackslash Cryptography}
  \item Reads the primary network adapter MAC address
  \item Computes \textbf{SHA-256 hash} of (MachineGuid + MAC)
  \item Posts registration request to \apipath{POST /api/v1/client/register} with:
    \begin{itemize}[label={-}]
      \item \texttt{fingerprint} — unique hardware identifier
      \item \texttt{deviceName} — Windows computer name
      \item \texttt{macAddress} — physical network address
      \item \texttt{ipAddress} — current LAN IP
    \end{itemize}
  \item Server creates a \texttt{PENDING} computer record
\end{enumerate}

The lock screen displays:
\begin{center}
\fbox{\texttt{PENDING REGISTRATION\\Fingerprint: a3f8bc12d5e7...}}
\end{center}

\subsection{Step 4: Admin Approves via Dashboard}

\begin{enumerate}[leftmargin=*, label=\textbf{\arabic*.}]
  \item On the admin machine, go to \textbf{Live Monitor} tab.
  \item An amber warning banner at the top shows: \textit{``Pending Workstation Enrollments (N)''}
  \item Click the pending workstation card.
  \item Enter:
    \begin{itemize}
      \item \textbf{PC Number} — e.g. \texttt{PC-A01}
      \item \textbf{Lab Zone} — Select from the dropdown (e.g. Turing Computing Lab)
    \end{itemize}
  \item Click \textbf{Approve \& Lock}.
\end{enumerate}

\subsection{Step 5: WebSocket Provisioning}

Immediately after admin approval:
\begin{enumerate}[leftmargin=*, label={\textcolor{aurxongreen}{\faCheckCircle}}]
  \item Server generates a unique cryptographic \texttt{qr\_seed} for this workstation.
  \item Updates database: status → \texttt{LOCKED}, saves \texttt{qr\_seed}.
  \item Pushes a WebSocket message to the connected workstation client:
    \begin{tcolorbox}[codebox, title={WebSocket Approval Payload}]
    \begin{lstlisting}[language=bash]
    {
      "type": "approved",
      "computerId": "uuid-here",
      "pcNumber": "PC-A01",
      "qrSeed": "randomsecretstring",
      "fallbackEnabled": true,
      "deviceName": "ALAMS-TURING-01"
    }
    \end{lstlisting}
    \end{tcolorbox}
  \item Client saves \texttt{computerId} and \texttt{serverUrl} to \texttt{config.json}.
  \item Lock screen transitions to active QR mode.
\end{enumerate}

\section{Physical QR Label}

After registration, place a physical label on each workstation:

\begin{tcolorbox}[alamsbox, title={\faTag\quad Physical Label Format (Per PC)}]
\begin{center}
\begin{tabular}{|c|}
\hline\\
\textbf{ALAMS Registered Workstation}\\[4pt]
PC Number: \textbf{PC-A01}\\
Lab: \textbf{Turing Computing Lab}\\
Device: \textbf{ALAMS-TURING-01}\\\\
\textit{Scan the on-screen QR code with your phone}\\[4pt]
\hline
\end{tabular}
\end{center}
\end{tcolorbox}

\begin{tcolorbox}[warningbox, title={\faExclamationTriangle\quad Important}]
Do \emph{not} print the dynamic QR code onto a label. The dynamic QR changes every 60 seconds and is cryptographically bound to a time window. Only the static identification label goes on the physical machine.
\end{tcolorbox}

\section{Verifying Registration Success}

On the Admin Dashboard → Live Monitor, the workstation card should show:
\begin{itemize}[label={\textcolor{aurxongreen}{\faCheckCircle}}]
  \item Status: \textbf{LOCKED} (amber badge)
  \item Client Agent: \textbf{ONLINE} (green)
  \item Watchdog Service: \textbf{ACTIVE} (green)
  \item Last Contact: within the last 15 seconds
\end{itemize}
